A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface

### 2021-10-25 sections/conclusion/main.tex
Add introduction and conclusion sections (#16)
* Update author names

* Update bibliography

Add papers that combine game theory and queueing theory

* Add initial draft of introduction

- First line (hook)
- Literature review methodology driven (game theory - queueing theory)
- Establish research problem
- Research objectives
- Map out the paper

* Cite papers on networks with tandem queues

* Remove unused citation

* Add some additional references

* Minor changes

- Remove unnecessary comments
- Apply suggestions based on alex and proselint
- Reference a paper on ED pressures
- Reference p

This study explores a 3-player game theoretic model between the decision makers 
of two queueing systems and a service that distributes individuals to these two
systems (section \ref{sec:model_overview}).
It also describes the construction of the underlying queueing theoretic model 
that has a tandem buffer and a single service centre (section 
\ref{sec:queueing_model}).
Furthermore, the formulas for the performance measures of the queueing model 
are also derived (sections \ref{sec:waiting_time}, \ref{sec:blocking_time}, 
\ref{sec:proportion_within_target}). 
The game theoretic model is then applied to a healthcare scenario by looking at
the interface between the EDs and the EMS (section 
\ref{sec:ems_ed_application}).
The inefficiencies that emerge from the perspective of the EMS were explored 
along with ways to apply some incentive mechanisms to improve them.
The main takeaways from this paper that were observed when playing the game
between two EDs and the EMS (section \ref{sec:application}) are:
\begin{itemize}
    \item Inefficiencies can be learned and emerge naturally
    \item Targeted incentivisation of behaviours can help escape inefficiencies
\end{itemize}
The former relates to the results of asymmetric replicator dynamics that showed 
that inefficient scenarios can arise by playing the game while the latter 
implies that the learned inefficiencies can be escaped by carefully applying 
certain incentives to the players.

This work also has certain limitations that need to addressed.
The model presented here only assumes the presence of two players that can
receive individuals. 
However, in a realistic healthcare scenario an ambulance may have to decide 
among multiple EDs.
An immediate extension of this work would be to consider a multiplayer system
that could represent a group of hospitals in a concentrated area.
Additionally, the game theoretic model that was created uses a discrete 
strategy space for the EDs (something that is also present in various pieces of 
literature~\cite{deo2011centralized, knight2017measuring}).
The single threshold parameter that is used for the ED's decision may not be 
the best way to describe the model.
In reality ED managers could have far more complex parameters for their 
decision making process.
Finally this work assumes that ambulances and EDs act in a selfish and rational
way by only aiming to satisfy their own objectives.
In fact, modelling behaviour in a healthcare setting is a much more complex 
task where some cooperation is often observed.
Another extension would be to explore the behaviour of the ED staff via an 
agent-based model. 
This in turn can be used to model ED staff as agents each with their own 
behavioural traits.

>>>>

This study explores a generic 3-player game theoretic model between the 
decision makers of two queueing systems and a service that distributes 
individuals to these two systems (section \ref{sec:model_overview}).
It also describes the construction of the underlying queueing theoretic model 
that has a tandem buffer and a single service centre (section 
\ref{sec:queueing_model}).
Furthermore, the formulas for the performance measures of the queueing model 
are also derived (sections \ref{sec:waiting_time}, \ref{sec:blocking_time}, 
\ref{sec:proportion_within_target}). 
This novel queuing model is the first contribution of the paper.
The game theoretic model is then applied to a healthcare scenario by looking at
the interface between the EDs and the EMS (section 
\ref{sec:ems_ed_application}).
The inefficiencies that emerge from the perspective of the EMS were explored 
along with ways to apply some incentive mechanisms to improve them.
The main takeaways from this paper that were observed when playing the game
between two EDs and the EMS (section \ref{sec:application}) are:
\begin{itemize}
    \item Inefficiencies can be learned and emerge naturally
    \item Targeted incentivisation of behaviours can help escape inefficiencies
\end{itemize}
The former relates to the results of asymmetric replicator dynamics that showed 
that inefficient scenarios can arise by playing the game while the latter 
implies that the learned inefficiencies can be escaped by carefully applying 
certain incentives to the players.
This applied game theoretic model is the second main contribution of this paper.

The model presented here assumes the presence of two players that can receive 
individuals. 
However, in a realistic healthcare scenario an ambulance may have to decide 
among multiple EDs.
An immediate extension of this work would be to consider a multiplayer system
that could represent a group of hospitals in a concentrated area.
Additionally, the game theoretic model that was created uses a discrete 
strategy space for the EDs (something that is also present in various pieces of 
literature~\cite{deo2011centralized, knight2017measuring}).
The single threshold parameter that is used for the ED's decision may not be 
the best way to describe the model.
In reality ED managers could have far more complex parameters for their 
decision making process.
Finally this work assumes that ambulances and EDs act in a selfish and rational
way by only aiming to satisfy their own objectives.
In fact, modelling behaviour in a healthcare setting is a much more complex 
task where some cooperation is often observed.
Another extension would be to explore the behaviour of the ED staff via an 
agent-based model. 
This in turn can be used to model ED staff as agents each with their own 
behavioural traits.

### 2021-10-25 sections/introduction/main.tex
Add introduction and conclusion sections (#16)
* Update author names

* Update bibliography

Add papers that combine game theory and queueing theory

* Add initial draft of introduction

- First line (hook)
- Literature review methodology driven (game theory - queueing theory)
- Establish research problem
- Research objectives
- Map out the paper

* Cite papers on networks with tandem queues

* Remove unused citation

* Add some additional references

* Minor changes

- Remove unnecessary comments
- Apply suggestions based on alex and proselint
- Reference a paper on ED pressures
- Reference p

A number of papers have been published that touch upon the use of 
queueing models together with game theoretic concepts.
In~\cite{FirmCompetition} the authors study a simultaneous price competition 
between two firms that are modelled as two distinct queueing systems with a 
fixed capacity and a combined arrival rate.
They calculate the Nash equilibrium both for identical and heterogeneous firms
and show that for the former a pure N.E. always exist and for the latter a 
unique equilibrium exists where only one firm operates.
The authors have also extended their model in~\cite{FirmCompetition2} by 
allowing the players (firms) to choose capacities. 
A main result from this paper was that equilibria, where both firms operate
together, are not socially optimal while when both firms operate as a monopoly 
they are.
Another extension of~\cite{FirmCompetition} was introduced 
in~\cite{FirmCompetitionExtension} where a long-run version of the competition 
was considered that also had capacity as a decision variable.
In~\cite{knight2017measuring} a normal form game is built that is informed by a 
two-dimensional Markov chain in order to model interactions between critical
care units.
An additional paper that focuses on competition is~\cite{fan2009short} where
the authors created a competition between two sellers where seller 1 supplies 
a product instantly and seller 2 is modelled as a make-to-order M/M/1 queue.
The game that is played requires the two sellers to make a choice on the price 
of the product and then seller 2 to set a capacity that guarantees a maximum 
expected delay.
In the above models, the players are attempting to increase their share of 
individuals choosing to queue.
In public healthcare type settings~\cite{knight2013selfish}, this is not 
necessarily the case. 
Players might not necessarily aim to maximise the so called `market share'.
Rather, the overall service needs to be considered as players aim to minimise
the overall congestion.
In~\cite{sadat2015can} a healthcare application was studied where patients 
could choose between two hospitals, where a utility function is derived that is
based on patients' perceived quality of life.
How individual behaviour based on congestion impacts the efficiency of public 
services is investigated in~\cite{knight2013selfish}.
The authors place the individuals' choices between different public services 
within the formulation of routing games and measure inefficiencies using a 
concept known as the price of anarchy (PoA)~\cite{koutsoupias1999worst}.
They show that the price of anarchy increases with worth of service and that is
low for systems with insufficient capacities.
In~\cite{deo2011centralized} the authors study the network effect of ambulance 
diversion by proposing a non-cooperative game between two EDs that are modelled
as a queueing network.
Each ED's objective is to minimise its own waiting time and chooses a diversion
threshold based on the patients it has.
In equilibrium both EDs choose to divert ambulances in order to avoid getting
arrivals from the other ED.
In this paper this concept is extended by allowing the ambulance service to 
decide how to distribute its patients among the two EDs.

@@

The EDs in the United Kingdom have to follow some set of regulations imposed to 
them by the National Health Service (NHS).
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.
This paper touches upon the following new aspects:
\begin{itemize}
    \item The emergence of gaming behaviour between EDs and the EMS.
    \item A queueing model with 2 consecutive waiting spaces where one would 
    serve as a parking space for the ambulances
\end{itemize}
The focus of this research is the construction of a 3-player game theoretic 
model between two queueing systems and a service that distributes individuals
to them. 
The resultant model will then be used to explore the emergent dynamics between 
the three players.
While giving some consideration to equilibrium behaviour similarly to the 
work of~\cite{FirmCompetition, FirmCompetition2, deo2011centralized, 
veltman2005equilibrium}, emergent behaviour is more precisely addressed by 
considering a number of learning algorithms.
This study explores two new concepts; getting performance measures for a new
queueing theoretic model with a tandem buffer and a single service centre and 
using a learning algorithm to model the emergence of behaviour.
The model is then applied to the healthcare scenario between two EDs and the 
EMS while looking into the inefficiencies that emerge and ways to apply some 
incentive mechanisms to improve them.
The EDs are modelled as two queueing systems each with a tandem buffer and a 
service centre. 
The performance measures are then used as the utilities of the game.
The novelty of the queueing model here is a contribution to not only the game 
theoretic literature but also to the queueing theoretic literature.
To the authors knowledge, no such model of a tandem queueing model controlled 
by a strategic decision has been considered. 
% TODO Be sure if this is correct

>>>>

A number of papers have been published that touch upon the use of 
queueing models together with game theoretic concepts.
In~\cite{FirmCompetition} the authors study a simultaneous price competition 
between two firms that are modelled as two distinct queueing systems with a 
fixed capacity and a combined arrival rate.
They calculate the Nash equilibrium both for identical and heterogeneous firms
and show that for the former a pure N.E. always exist and for the latter a 
unique equilibrium exists where only one firm operates.
The authors have also extended their model in~\cite{FirmCompetition2} by 
allowing the players (firms) to choose capacities. 
A main result from this paper was that equilibria, where both firms operate
together, are not socially optimal while when both firms operate as a monopoly 
they are.
Another extension of~\cite{FirmCompetition} was introduced 
in~\cite{FirmCompetitionExtension} where a long-run version of the competition 
was considered that also had capacity as a decision variable.
In~\cite{knight2017measuring} a normal form game is built that is informed by a 
two-dimensional Markov chain in order to model interactions between critical
care units.
An additional paper that focuses on competition is~\cite{fan2009short} where
the authors created a competition between two sellers where seller 1 supplies 
a product instantly and seller 2 is modelled as a make-to-order M/M/1 queue.
The game that is played requires the two sellers to make a choice on the price 
of the product and then seller 2 to set a capacity that guarantees a maximum 
expected delay.
In the above models, the players are attempting to increase their share of 
individuals choosing to queue.
In public healthcare type settings~\cite{knight2013selfish}, this is not 
necessarily the case. 
Players might not necessarily aim to maximise the so called `market share'.
Rather, the overall service needs to be considered as players aim to minimise
their experienced congestion.
In~\cite{sadat2015can} a healthcare application was studied where patients 
could choose between two hospitals, where a utility function is derived that is
based on patients' perceived quality of life.
How individual behaviour based on congestion impacts the efficiency of public 
services is investigated in~\cite{knight2013selfish}.
The authors place the individuals' choices between different public services 
within the formulation of routing games and measure inefficiencies using a 
concept known as the price of anarchy (PoA)~\cite{koutsoupias1999worst}.
They show that the price of anarchy increases with worth of service and that is
low for systems with insufficient capacities.
In~\cite{deo2011centralized} the authors study the network effect of ambulance 
diversion by proposing a non-cooperative game between two EDs that are modelled
as a queueing network.
Each ED's objective is to minimise its own waiting time and chooses a diversion
threshold based on the patients it has.
In equilibrium both EDs choose to divert ambulances in order to avoid getting
arrivals from the other ED.
In this paper this concept is extended by allowing the ambulance service to 
decide how to distribute its patients among the two EDs.

@@

The EDs in the United Kingdom have to follow some set of regulations imposed to 
them by the National Health Service (NHS).
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.
This paper touches upon the following new aspects:
\begin{itemize}
    \item The emergence of gaming behaviour between EDs and the EMS.
    \item A queueing model with 2 consecutive waiting spaces where one would 
    serve as a parking space for the ambulances
\end{itemize}
The focus of this research is the construction of a 3-player game theoretic 
model between two queueing systems and a service that distributes individuals
to them. 
The resultant model will then be used to explore the emergent dynamics between 
the three players.
While giving some consideration to equilibrium behaviour similarly to the 
work of~\cite{FirmCompetition, FirmCompetition2, deo2011centralized, 
veltman2005equilibrium}, emergent behaviour is more precisely addressed by 
considering a number of learning algorithms.
This study explores two new concepts; getting performance measures for a new
queueing theoretic model with a tandem buffer and a single service centre and 
using a learning algorithm to model the emergence of behaviour.
The model is then applied to the healthcare scenario between two EDs and the 
EMS while looking into the inefficiencies that emerge and ways to apply some 
incentive mechanisms to improve them.
The EDs are modelled as two queueing systems each with a tandem buffer and a 
service centre. 
The performance measures are then used as the utilities of the game.
The novelty of the queueing model here is a contribution to not only the game 
theoretic literature but also to the queueing theoretic literature.
To the authors knowledge, no such model of a tandem queueing model with a pair 
of parameters for the buffer has been considered.
% TODO Be sure if this is correct
